Refinements for open automata

### 2022-07-13 2021-Refinement/paper/refinementarticle.tex
typos + comment under Def 6

The notion of refinement  captures the relation between  a specification and an implementation of the same component. Refinement entails that one system can be considered as a more precise version of the specification, featuring all the specified behaviours with more concrete details. 
 It is usually defined as trace inclusion or simulation  \cite{Milner:1980,Kouchnarenko:2007}; this ensures that all behaviours of the specification must belong to the behaviours of the implementation . 
 This definition, which is based on systems whose behaviour is fully defined, is not well-suited for open systems,  as it requires to reason also about unspecified behaviours.

@@

We thus define the following property that can be understood as follows. Take a specification automaton and its implementation. Consider two related states for the refinement relation. Suppose an identical transition exist in both automata (with same guards and same label), then either the reached states $s_1'$ and $s_2'$ are related by refinement, or there exists another transition of the specification that is also identical and leads to a state $s_1''$ that is related to $s_2'$ in the refinement relation.
Finally, we should take into account that 1) instead of equality, we should check if there is a state such that both transitions are feasible (the conjunction of guards is satisfiable), and that  2)instead of finding another transition of the specification we might find a family of transitions that cover all the cases of the transition of the implementation. This leads us to the following definition.
\begin{definition}[Preservation along identical transitions]\label{def:preservation} \\
Let \(\OAg[1]\) and \(\OAg[2]\) be two OAs.
A relation \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) between  \(\OAg[1]\) and \(\OAg[2]\)
  is preserved along identical transitions if  it satisfies the following:
% TODO: isn't it a2 = a1x instead of a1x = a1
% TODO: ANS:  we have a1 = a2 so whatever, it is the same, we can change if it is more ``natural''
\begin{multline*}
\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} \mpar{t_1, t_2} = \mpar{\OTx{1}{}{1}{1}, \OTx{2}{}{2}{2}} \in \fOT{s_1} \times \fOT{s_2} \\ 
\bigsymb{\bigexists} 
\mpar{t_{1x} = \OTx{1}{x}{1x}{1x} \in \fOT{s_1}}^{x \in X},
 V_1 \uplus V_2 \uplus \biguplus_{x\in X}\fvars{t_{1x}} \uplus \fvars{t_2} \uplus \fvars{t_1} \vdash \hspace{6cm}\\  R\mpar{s_1, s_2} \wedge 	g_1 \wedge g_2 \wedge \alpha_1 = \alpha_2 
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} 
\implies \\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{x\in X}}
\mpar{
\alpha_1=\alpha_{1 x}  \land J'_{1x}= J'_2 \land \bigwedge_\subbox{j \in J'_2 } \beta_{2j} = \beta_{1xj}\land g_{1 x}\land
 R\mpar{s'_{1 x}, s'_2}\psubst{\psi_{1 x} \uplus \psi_2}}\\
\end{multline*}
\end{definition}

@@

Note in this definition, instead of matching an instantiated transition (with all variables assigned) of the first automata to another instantiated transition of the second, it matches an open transition $t_1$ to a family of covering open transitions $t_{2x}^{x\in X}$. %Note also the second part of the conjunction of condition (2) expresses the deadlock reduction between automata (as introduced in Definition \ref{def:dpwd}).

@@

Yet according to the refinement  relation between   $A_2$ and $A_3$ the following is satisfied:
\begin{multline*}
\bigforall \mpar{s_{2p}, s_3} \in S_1 \times S_3, \bigsymb{\forall} \mpar{t_{2p}, t_3} = \mpar{\OTx{2p}{}{2p}{2p}, \OTx{3}{}{3}{3}} \in \fOT{s_{2p}} \times \fOT{s_3} \\ 
\bigsymb{\bigexists} 
\mpar{t_{2px} = \OTx{2p}{x}{2px}{2px} \in \fOT{s_{2p}}}^{y \in Y},
V_2 \uplus V_3\uplus \biguplus_{x\in X}\fvars{t_{2px}} \uplus \fvars{t_3} \uplus \fvars{t_{2p}} \vdash \hspace{6cm}\\  R_{23}\mpar{s_2, s_3} \wedge 	g_{2p}\wedge g_3 \wedge \alpha_2 = \alpha_3 
%\wedge \bigwedge_\subbox{j \in J'_2 \cap J'_3} \beta_{2pj} = \beta_{3j} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_2}. \\
\bigsymb{\bigwedge_{x\in X}}
\mpar{
\alpha_{2p}=\alpha_{2p x}  \land J'_{2px}= J'_3 \land \bigwedge_\subbox{j \in J'_3 } \beta_{3j} = \beta_{2xj}\land g_{2p x}\land
 R_{23}\mpar{s'_{2p x}, s'_3}\psubst{\psi_{2p x} \uplus \psi_3}}\\
\end{multline*}

For any open transition $t_{2px}$ originating from $s_{2p}$ we have 
the following by definition of refinement between  $A_2$ and $A_3$:

>>>>

The notion of refinement  captures the relation between  a specification and an implementation of the same component. Refinement entails that one system can be considered as a more precise version of the specification, featuring all the specified behaviours with more concrete details. 
 It is usually defined as trace inclusion or simulation  \cite{Milner:1980,Kouchnarenko:2007}; this ensures that all behaviours of the specification must belong to the behaviours of the implementation. 
 This definition, which is based on systems whose behaviour is fully defined, is not well-suited for open systems,  as it requires to reason also about unspecified behaviours.

@@

We thus define the following property that can be understood as follows. Take a specification automaton and its implementation. Consider two related states for the refinement relation. Suppose an identical transition exist in both automata (with same guards and same label), then either the reached states $s_1'$ and $s_2'$ are related by refinement, or there exists another transition of the specification that is also identical and leads to a state $s_1''$ that is related to $s_2'$ in the refinement relation.
Finally, we should take into account that 1) instead of equality, we should check if there is a state such that both transitions are feasible (the conjunction of guards is satisfiable), and that  2) instead of finding another transition of the specification we might find a family of transitions that cover all the cases of the transition of the implementation. This leads us to the following definition.
\begin{definition}[Preservation along identical transitions]\label{def:preservation} \\
Let \(\OAg[1]\) and \(\OAg[2]\) be two OAs.
A relation \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) between  \(\OAg[1]\) and \(\OAg[2]\)
  is preserved along identical transitions if  it satisfies the following:
% TODO: isn't it a2 = a1x instead of a1x = a1
% TODO: ANS:  we have a1 = a2 so whatever, it is the same, we can change if it is more ``natural''
\begin{multline*}
\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} \mpar{t_1, t_2} = \mpar{\OTx{1}{}{1}{1}, \OTx{2}{}{2}{2}} \in \fOT{s_1} \times \fOT{s_2} \\ 
\bigsymb{\bigexists} 
\mpar{t_{1x} = \OTx{1}{x}{1x}{1x} \in \fOT{s_1}}^{x \in X},
 V_1 \uplus V_2 \uplus \biguplus_{x\in X}\fvars{t_{1x}} \uplus \fvars{t_2} \uplus \fvars{t_1} \vdash \hspace{6cm}\\  R\mpar{s_1, s_2} \wedge 	g_1 \wedge g_2 \wedge \alpha_1 = \alpha_2 
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} 
\implies \\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{x\in X}}
\mpar{
\alpha_1=\alpha_{1 x}  \land J'_{1x}= J'_2 \land \bigwedge_\subbox{j \in J'_2 } \beta_{2j} = \beta_{1xj}\land g_{1 x}\land
 R\mpar{s'_{1 x}, s'_2}\psubst{\psi_{1 x} \uplus \psi_2}}\\
\end{multline*}
\end{definition}

@@

Note in this definition, instead of matching an instantiated transition (with all variables assigned) of the first automata to another instantiated transition of the second, it matches an open transition $t_1$ to a family of covering open transitions $t_{2x}^{x\in X}$. %Note also the second part of the conjunction of condition (2) expresses the deadlock reduction between automata (as introduced in Definition \ref{def:dpwd}).   
\TODO{The first and second conditions coincide with the natural way to prove inductively that an automaton simulates another by starting with the initial state. The third condition ensures that the refinement relation does prevent the introduction of deadlocks.}

@@

Yet according to the refinement  relation between   $A_2$ and $A_3$ the following is satisfied:
\begin{multline*}
\bigforall \mpar{s_{2p}, s_3} \in S_1 \times S_3, \bigsymb{\forall} \mpar{t_{2p}, t_3} = \mpar{\OTx{2p}{}{2p}{2p}, \OTx{3}{}{3}{3}} \in \fOT{s_{2p}} \times \fOT{s_3} \\ 
\bigsymb{\bigexists} 
\mpar{t_{2px} = \OTx{2p}{x}{2px}{2px} \in \fOT{s_{2p}}}^{y \in Y},
V_2 \uplus V_3\uplus \biguplus_{x\in X}\fvars{t_{2px}} \uplus \fvars{t_3} \uplus \fvars{t_{2p}} \vdash \hspace{6cm}\\  R_{23}\mpar{s_2, s_3} \wedge 	g_{2p}\wedge g_3 \wedge \alpha_2 = \alpha_3 
%\wedge \bigwedge_\subbox{j \in J'_2 \cap J'_3} \beta_{2pj} = \beta_{3j} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_2}. \\
\bigsymb{\bigwedge_{x\in X}}
\mpar{
\alpha_{2p}=\alpha_{2p x}  \land J'_{2px}= J'_3 \land \bigwedge_\subbox{j \in J'_3 } \beta_{3j} = \beta_{2xj}\land g_{2p x}\land
 R_{23}\mpar{s'_{2p x}, s'_3}\psubst{\psi_{2p x} \uplus \psi_3}}\\
\end{multline*}
For any open transition $t_{2px}$ originating from $s_{2p}$ we have 
the following by definition of refinement between  $A_2$ and $A_3$:
